Ouch!

YEah, all right, so... Ahm, Chris, culd you tell me a little bit about how you got started in aztec music or. ... mesoamerican musical instruments

Sure, I was actually ahm, there was a group called Ancient groups, and ahm, ahm,
the guy at the time, the musical director at the tiem asked me to---
a duo thing with him in Pasadena, 

And I didnt know him at all
The person who normally worked with him couldnt work with him
and said "I told this guy to call you"
-i said ok
so I talked to this guy
and he said "just show up and we're gonna do this presentation"
I said "ah, ok"

So he played multiple instruments from, mostly east europeans instruments
so ahm, I did that and then he said "That was it!"
and he said "I now you play instruments of india"
and I may call you in the future
I said Ok

Years, literally years goes by, and he said "I'm playing with this group and I'd like you to come and Id' like yo too play indian instruments"
I said ok
So it was malik soul who was african, african?¨Martin Espino, and this guy Joghn. John was the guy that asked me
So, my Van was in the shop so I called jhon and I said, and it was in Ventura,  I said "John I have no way to get in there, so I... If i can get a ride tats great if not, i cant do it"
 he said im gonna ask martin too get you
 i said ok
 so martin espino, thats when i met martin
 So he picked up, I've been here almost 20 years this spot, so he picked my indian instruments when we went u there and he starred at me and it turnes out, we knew a lot of the same people, but we have never been mentioned to anybody else.  So anyway, so I did that thing with them and then Martin called me like a month later and said "you know= I do this things for children and i was wondering if you could play drums" and -i said "my indian drums?"  he goies, no, no, "my drums from india?" and he said "no you would play my instruments, my ahm..." He didn't have a huehuetl at the time, he just had water drums he said "you would just be playing those" I was like ok. So i went and I... we didn't do anything danza related, there were no danza rhytms, cause I didn't know any danza rhyutms and at that point he didn't either.

 So, he said "ah, bring what you know into this" all right? So I, you know, I played drums for years, pleayed indian music, north and south indian classical an folklore for years, riht? and i didnt know this instruments, so i started working with martin and we started making compositions and... the thought process was there's more to that music that just the basic pawls, and to me it's like... 

 you know there's dance music, there psychodelic music, there's reigiois music and there's on the lack of a better word art music. Now any dance music of any tradition is repetitive and the same, so I could see to point why that is whats left, cause i cant think of any other population that has lost 90& of their population and mantain what.... of their culture was, or to that point, as for example my indian teachers or my african teachers they did not loose 90% of their populations and they argue all the time about their tradition and what the tradition is and what it shouldnt be and what they think it should be and what it is and etc etc etc 

 so... i't like just because we dont' have any audio of what it sounded like other than what it has existed thorough danza, does not mean that tehre wasnt other things going on, and its like... to me, that the pyramids, well number one there's a thousand pyramids in mesoamerica and there's  a hundred in asia.

 And the pyramids are built in relation to the cosmos and the stars and the universe and time lines and you know etc, etc, etc. So to me, to think that they were just doing this is... ahm... I can't be-, I can not conceive that that is all they did. Im sure that there was many more things but we dont have any audio

 SO i never say ever never ever ever say, if i do a lecture, i say "ok, we're not gonna say rehispanic or not gonna say precolombian or  not gonna say precortesian, were not gona say those words" 

 and i have to say my wife is from, my wife is from La valle del yaqui form sonora mexico and ahm the reason I was saying that Im like

 So i say "were not gonna say this words for this lecture", we're gonna call them indigenous instruments of mexico and mesoamerica and now i'll tell you why, because people north of the border or what is now a burder, if you see an apache playing drum or a salika playing a flute or an coyek playin and you ask them what theyre playing they say im playing "a Krsna flute or an apache drum" They dont say "im playing a predutch flute or Im playing a pregerman drum " that only existed in mexico

 Not even in south america do they call it "precolombian" or "precortesian" they dont, they call it by what it..like theTaino, the  Taino flut is the tayno flute, they don't say its prespanish, 
 so this whole concept to me is te colonization of the mind, which is which has been propagand and the people who argue with methe most about it than anybody else are people from MExico City and they say "that's what it is" and I'll say well my father is Garcia and my mother is Bello, but I would ever call my motehr Pregarcia, you have to call things for what they are

 So as I started playing this instruments, i started doing research on this instruments and going back and just doing my own research because i wasnt playing anything danza related at all, the rhytms were all coming upwith these rhytms that had nothing to do with danza, so i met danza drummers ah... martin gardo, who is the one of my favorite danza drummers and his teachers Eric Barraza, Octavio Alere more and Diego castañeda who are Don lazaro Calpulli, and ahm... they started inviting me to play with them on Depo, so I would play depo with them when they played to

 So i learned the danza but we never incorporated that into what we were doing cause there was a, a .. at that time, for whatever reason, martin was not working with danzantes at all, in any way they perform, and it was, it was ah... which is fine, to me it was like ah, no we would go like this, this is the history of mexico, this is who we are now, and this is cortes, but I mean it goes back to hear and also, ah.. we're not just taking this, you lnow? We're playing for other people, we were playing yaqui, water drums, we're playing tepejuano, boud instruments

 We're taking from different indigenous cultures and coming up with new things, rigt? So i never ever said, and I don't think he ever say that we were doing ah traditional music

 And i wouldnt never say doing thateven if somebody showed me a chant, like i know chants but i never performed them because i was speaking to ah... ah... i dont even know the terminology, a native american elder meaning nirth of the border and he was saying the problem that he had is that there was a lot of young people that are learning chants from youtube that are ah from traditional ceremonies that are not aware that that chant it was taught to specific family and yourspecific family was allowed to do that chant and that was was a very specificidea, the whole thing, so they dont know the words theyre making up to for the chants.

 So if i'm with somebody, like i was with luis, and I work with Rayman Aramirador who's otomie,when they showed me that stuff I mean i accompanied them, they showed me what they were doing, but i would never do that withot them just because i dont 

 And i, when i went to ah study, had a full scholarhip to study indian musci at cal arts, so at that point in time I was the only hispanic, i was one of three hispanics in cal arts from 5hundred students and 80%of the people were not from california. So when I said "indian music" Im not indian so i never dressed like an indian

 So my indian teacher, that's him up there, God bless him. he was totally fine, uou didnt have to be indian for him to teach you. So I.. that's why i dont wear, i dont wear ... out of respect of the people that wore .... So they know that, so when they ask me to play with them, they know im not gonna be like this, they dit have a roblem thats why they ask me like if they had a problem, they wont invite me, because they know im doing it out of respect to them ahm

 So that's how I got into it and I started studying, started playing like traditional danzantes, but then i met kuis Perez who is much more ahm... Luis is the only persone I, no, actually no, the only human being that speaks 6 indigenous languages that for the last 40 yeas has traveled and lived with indigenous people, learned to make their instruments, learned theyr songs, learnd their chants, and ah... Work with them, they're doing a documentary on it, on the last 5 years, this is gonna come out next year

 So this guy is very, he's informed, and he's knowledgable, right? Information, knlowedge wisdom is the last thing, right?,  but hes nformed and hes very knowledgable. Much more so than anyone ellse I have, that I have actually met, ah

 and he would came to UCLA in 1980, he was invited to UCLA to do the first symposium on Mesoamerican instruments, 1980, so that's when ahm... people first saw that aount of instruments. I mean he actually has 2 thousand ----  instruments that are alone like he's like a living treator from mexico as well as replicas, so he has all of these instruments, and knows all of these different vernacular languages because all the people are different, right? So he tedns to either, ah, he might do a.. So when we performed, he might walk out and speak in raramuri and then will do something in yaqui or will do something form neceit, so he's going... He's growing from all these places, right? so he never says "Im doing traditional music" He never says "Im doing precolombian, or prehispanic or precortesian" which is that's... well actualy he says precolombian (laughs). But ahm, and I never say "conquistador", I say because 2You knw? nobody in japan, nobody in japan calls us conquers. (yeah!).. OR germany, I said in MExico is the only people who calls it people who invaded tem conquers. There's no other country in the wordl

 IF you ask them, if you ask a spaniard I¿ve been in 28 countries in 5 continents,ifyou ask a spaniard about the Morse they sat thet invaded us, they dont say they conquered us. So this whole mindset to me is, ah, is just a mindset .

 But anyway, I got into this drumming through the danzantes, to learn what they were doing but i never wanted to replicate it i just wanted to understand how it worked, how the phones worked, and ah, theres actually stuff on Filmwor where we're giving a lecture, ah, which I have where ah... *laughs*, I mentioend at this lecture ah, that, the drummers following the dancers, the dancers are not following the drummers. And ah, it was ah, we got into actual verbal exchange, it was like "no, no, no, the dancers are following the drummers!", i said "no, no, no, the drummers..." but this is actually on film i can send that, but that was whith martin (yeah!) cause I was like no.. cause he, at that point, he wasnt invovled with danzantes.

 ah, and so the whole time im playing drums, I'm still playing rumba, im still playing indian music, and im playing indigenous instruments and doing indigenous music
 So the problem i have are that i see now, with instrument makers as an expamle
 if you, like all the guys, the three guys that are making instruments, they will thell you very specifically that theyre tunning the instruments pentatonically and I'll say that te instruments are not tuned pentatonically and ill send them the fourty flutes from the anthropological museum of the tunies and none of them are pentatonic, and they're like, "yeah, but its easier for people to play" and I go "thats fine, but you need when you make a flute and you sell it you need to tell the person that its not... that" Right?To me its just like, i dont even understand that

 I mean i understand somebody wanting to play because they love, they like the sound, i get it, but i mean that's not what it was.

 Where do you think thats coming from...?

Where do i think thats coming from? Well, I think it's... from them wanting to generate income. And you know, that people like what they look like and they like what they sound like,and it's easierplay pentatonic, i mean i understand it, i dont agree with it  totally understand it i get it, you know? but you need, that needs to be, ah.. the people need to be informed when buying stuff like that
you know?
ive bee to things where
there was this guy, ahm, he had made a, he had brought an aztec calendar which was not a calendar and he was explaining the calendar to this people and they were giving him money and he was dressed like in a, like a dodgee, u know? like an indian, like an indian india, and he's sitting there and while im listening to this guy going on and on about this, and the you know? the people leave and then when they leave he's sitting there and ii said to him "you know none of what you said is what it is?" and he said "i know, but thats what they wanna hear" that's really sad... you know? Thats really sad to me

thats not representativ of our culture, of our people, of the indigenous people, its not cool.. but, 

Is this person still in circulation or?

you know? i only saw ihm that one day, but im sure he is. (yeah) but ahm...

so what i did was, ah, i went to england, i was in england and there's a gentlleman named Ian MErsall, do you know, are you aware of him?

i dont think ive known him

Ian mersall has a website called mexicanlord.uk.com and hiswife is from Mexico city and he's from england, and he met me in london and he said "can you make some videos and demonstrating instruments?" and i did, theres 80 thousand people that has seen them now 80 thousand people. Ahm, but the only tag on the, youknow when you put it on youtube, they put a tag, the only tag is the name of the instrument, so i didnt' put "slit drum2 i didnt put my name, so teponaztli, teponaztli, huehuetl, so number one that means the person need to know what they're looking for and there's 80 thousand people dont know what theyre looking for, they only been up to 3 years (yeah!). So that's kind of interesting to me...

So working with Luis Perez xixoneshtli, he has a... I could send it to you. He has a, a, so he's, he's a.... So like, you know martin espino has a degree in classical guitar. Luis perez has a degree on classical guitar and piano, like Javier Key he doesn't have a musical education 

so these guys have a different erspective on music on general, so Luis when we do things he'll be ah, I mean... a lot of times with them, when Im working with somebody i just say "just play the rhythm" let me film you and record you, and that's what im based on. So we don't write anything down, just like all of my indian stuff nothings written down its all oral, so I mean just play it forme and let me see ow you doing it let me see how your hands are working and thats what ill do, because if you notice in any folkloric tradition, wheter is chinesse japanese african mexican indigenous, if youre playing a drum, they play hand to hand. You know? theyre gonna do ah, you know' they're playing ah, [drums playing]

They're not doing [drums playing]

They're not doing that, everythings hand to hand. Everythings hand to hand

So, the choruses, somebody playing with sticks... the choruses is nobody wearing chachiates, not one person. There's only one chodes that shows a person wearing chacheotes,its the dressing code in the mayan. The oldest image of the tlapanebl is Chiaas from Morelos and is not mexica.

So, so, remember i was saying if you ask somebody when did the mexica first came into the valley in mxico? when was the triplealliance? whatwas the triple alliance about? No clue! So in india.... I have four teachers, that I studied with. And they will say study with him, right? they can give you stuff that I... it's another perspective of me.

But here, the oral tradition is like "this is my maestro and i do what my maestro says" which is, if you just get this out of context is very simple, right? if youre in grammar school and you haveone teacher for all six grades, that could be done, but if you have a different teacher youd have different perspectives. But unfortinately what happens in this tradition is "hatever is said by the maestro is considered what it is" so when you speak to other contemporaries, like, Heather he's a very different perspectivie on anything, you could ask him about anything that you've documented he¿ll oure gonna get a very different perspective, cause  not only is he an elder but hes actually older, so he has a very different perspective on that stuff

So when I played this instruments i played them the way they played. so i dont use any, no i cannot use north indian or sout indians hand or finger techniques but if you read clavallero, yesue he says "that that instrument was played with their fingers and not with their hands and that each finger they could do what their fingers ...... dexterity. 

Right? Cause two sticks is to baters two hands are two beaters, your fingers you have tn beaters, El cantar este mexicano, have pieceswritten for one singer and ten clap on huehuetls, ten clap on uehuetsl and one singer and the pitch on that song is taken from the picture drawn so the person has a low voice.Soyoure not gonna have ten gays playing, i mean... you need to hear the singer

so i mean and ill send you some other stuff, i dont know if you had a chanc to see it but Motolini and toribio had their described how the danza was done compared o how its done now, that nobidy does anything thats writting, and like the concept of the four directions... is not there!

That doesnt mean that it wasnt done, but it doesnt mean that it was done. So now when you see, and i gave you the example of Ray Shankar, when Ray shankar first came to america, before you played the composition, you played the alab, the alab is  based on the **** without time, but it'sbased on, on, emutng all on tht based on deran and in india **** two hours before you play the piece, when he came to america he said "i coundt play longer than 8 minutes" cause people would lose ther concentration, so then he played the piece cause when the piece and the drummer comes in and its rhytmic....so not all  indian music is  presented that wAY. You have these ten minutes ... laps and then this hour long compositions, but its not the way it was done, but thats what it should become.

The same with danza, when I grew up in mexico city and i see the danzantes rehearsing theyall  were  wearing the same thing, they wear jeans a tshirt and a bandana all of them, every single one of them, whether im at bellas artes, whether im at the zocalo, whether im at the ah, i cant remember the name of the place, but.... they alldo the same thing. and they come into mexico city to rehearse.

and they dont wear regelia at all, at all, at all, at all, until they perform.

Right? Then everythings like its just so hard to travel that concept. They dont have,, they do not accept money, they do not put a thing on for money, they do not, not one of them. So all of this was very different to me, to experience that i was there for 8 days and i went there every day just to see these guys just to see what they were doing how tey were doing, somedays you will have 80 dancers and 8 drumers, somedays youll ave 6 drummers and 5 dancers at these groups, ah...

Erick Baraza, i dont know if you, areyou here next thursday'

Yeah, i think so, down at long beach, but I...

Erick baraza every tuesday runs a drumming, for lack of a better word ill say its a danza class, but its actually for the drummers. Its for the drummers because what happened is that they were finding that when they went to the ceremonies there was peple who were drumming that werent drummers or there were people that were drumming that didnt know, so for example when youre playing for danza, you have to play so tehy can dance, so if your playing a... Aguila, you know? {sounds} What you regarded the ceremony, they're be like (sound) and the dancers couldnt dance, they just couldnt.... And then theyre just trying and theyre not, and they were losing, some of the drummers werelosing the whole perspective, so you sure mean that... you should mean that the whole thing

youre not serving yourself, and youre not serving the dancer you and the dancer are one, so your.. its making the whole thing.

So ah a lot of the more experienced druming stopped going to the ceremonies, because they were told to police the younger guys and they were like "i dont go ther to police uounger guys i went there to plau" so there was this whole thing, and so it happens where a lot of the guys I know, like ................ they dont' go because its like... and they are the guys! Theyve been i mean, how old is martin¡ I think hes 28 now, you know hes been druming since hes 14, but hes like... he just doesnt wanna be police man, you know?

so thats unfortunate, because then you start seen a lot of the

******
I know youll speak to Rim, Rim is very artic..., he's bery quiet, but hes very... he knows his stuff. One time we were doing a rehearsal and ah they were doing this dance, and we were working with four calpulis for this thing and he was ah, the designated leader, and ah, i cant remember which dance it was but ah, they did this one movement, three times or maybe four times, the fourth time and only the fourth time if it is four times, they... when they turned and spin the last time instead of just ah mark in there, their foot, theyre supposed to turn their, the bottom of the foot up to the heavens and drag their foot and leave a mark in the dirt, and so he was watching them dance and then, and this arepeaople that have been ddancing 15 years minimun, and hes been dancing like 40 or something, and hes been watching them and he sais, ah, why dont your guys ahm, why aren you guys dragging your foot?
and theyre like what? And hes like" why aren you guys dragging your foot?"
and they said "ell because, its foir times we never did it the first three times so we said why should we do it the fourt time?" and he stood up and explained for an hour what that one movement was about
youre representing the comet through the sky and thats why you drag your foot the last  time and he went into this long lentgy explanation of it so he is like... youre like, were all like so then when everybody left i said Ray! I said ah, how could you didn't... ah... say anything before, you know? about anything? and...cause hes just very quiet and he says "cause should sent you all day and tell them what they dont wanna know?" so thatsa very... 

my allusion india like that too is like they dont wanna know, I'm not just gonna speak and ... I've been.. from japanese elders, chinese elders, native elders,its the same thing, its like "I could be here talking all day" So its interesting whats changing,you know? and why its changing? and ah,

Also is interesting what drumers like noticed pretty much anyone can show, like I could almost buy a drum and then just show up and start druming like someones... the one where ive been before, noone will blink, even if I... not playing it correctly. ah, and Ive talked to some other, ah, like captains of ah... desk groups and they find it frustrating beyon the leave, you know? I was talking you about disciplina that you cant have disciplina because  ah specially in danza theres a sense that people will leave cause there's so many capullies, ah, you know? some of them will just say "then ill  just go be with there instead" people have it really, theyve been struggling tofigure how to get these worked out?

Right...
Well i think that just comes from a lack of information, I mean... "you dont know, you ask". I mean, this rule is basic to me is like, you would ask your teachers thats ok but if i go,like to certain teachers that will tell you ********************** and you have to do this

but the thing that's interesting to me,ive noticed the same thing in studying indian music,theres a lot of people who are not indian that are very strict on the rules of playing indian music taht are not indian, whereas indian would never do that, allthe indian,indian from india that ive known would never do that, but theres people that are not indian, fro other destinies and cultures who are very well in new dali is like this, you know? and ah, ive see that here too. You know? theres peple that are not mexico city the mexica people, right?

so if you hear from guerrero are you from sinaloa are you from acapulco, youre not mexica, youre not. Blood, youre not. Youre not. 
Just like, you know, like my wife is from La valle de yaqui, right? Now, let me get a context.
Theres a martian who loves the  beach sports, he know all the beach boys songs, he buys huaraches, he buys white shirts he buys shorts and he goes down to malimbu and he learns to say 20 things that the beach boy says, hes not a beach boy, hes never gonna be a beach boy, hes a martian, so its the same concept. Just because you coudlsay tlazucomatli or the, doesnt make you indigenous and most of the people are not from MExico, mexico city, you know? I... so,

The thing thats iportant to me about living and live is your intention why are you doing what you do, your intensity howoften you do it, your integrity,how you do it. Intention integriti intensity passion and balance.

And, when i see, you can see, you can go to any ceremony and see 80 dancers, and theres like... those four are dancers, the other 76 are here.. (laughs)...but you cannt tell that those 4 dance way more than the other 76 and its very apparent, you can just watch them

so when you read the, ah, spanish ah observations, where they what he sais, when they, when they bend their arm,their their, this was totally parallel to the ground, all of them, you know? this was not lie this, it was all like, it was so perfect,it  was beyond perfection and they say it was beyond perfection then nobody in europe and nobody in our dances could what they did, and I mentioned to you, i told the stew danzante and he was like "well if its perfect they couldn be indigenous" and im like... Just look at the pyramids, i cant even understand why anybody would think that

a danzante said that to me "if it was perfect it wouldn be an indigenous" and i was like I mean i understood what he said but im like i dont agreed at all cause the pyramids are not just like no "Put it over there Joe!" i mean they jsut didnt put it you know what i mean it was like where does it is, why it is, when it is is very important, you know such as there so

to me its just like no

i mean, right none of us are perfect. Never will be, but you strive for perfection. So, when I see 80 dancers and i see 4 of them are dancers, and I dont even, im not even talking about technical abilities im talking a bout their intentions, how they carry their sause, how they move, who they are, its very apparent who is who and who is not. I mean, some people are doing it as a matter of cultural pride which is fine, thats fine, but maybe just dress like that and not dance... you know? 

But that's my totally biased subjective opinion, but its like, you know?
That wouldnt happen in india.

You know? Nobody can just sit and play in india that would never happen, that would never happen. So what, what... anyway go ahead1

The commment of a perfection and this idea of... to be identity as well... mexicano vs indigena.... and danza.... that's very complicated as well, what has been you observation for you now when its been used for national pride ... identify as indigenous through dance performance?

Well I think there's a lot of lines, and theres a lot of divisions, so ahm... and they all exist and they coexist whether i agree with them or not, amm, but Ill give you an example of that.... I was playing a, tlapanohuehuetl, this is the tlapanohuehuels not the huehuetl, i was playing the tpalanohuehutl and ah, the maestro came up to me and said "do me a favor when you play this part of the rhytm i want you to play the stick on this side of the drum, set it on the rhym" So im playin with sticks for the first place, cause he plays with sticks and im just doing what hes asking me to do and that hes asking me to play this fun part of the rhytm you know, like (sounds) and I can do that, no problem. So im doing it and they take a break, take a break and then this danzante comes to me and says, can i ask you a uestion? Cause im obviously not in the.... im obviously not like that, not dressed like that and he say "why are you hitting the drum on the side" why am i hitting the drum on the side? He was like yeah!  "why are you doing that?" "is that traditional" that was the question "is that traditional'" and i said "ah, your maestro asked me to play that rhytms with the drum on the side, ive never done it that way" second answer "is this traditional?" my question was "have you seen any of the codices?" answer was no, and i said "well look at the codices, you will not see... theres three images of people laying with one stick and allthe other ones other people are playing with hands and the only people playing with sticks is the deili and then everybody else was playing with theirhands so is it traditioal?  id say no, laying with sticks isnt even a traditional, but you have velcro,youre wearing a bucle, a bucle is not traditional, youre wearing eathers that are not even indigenous to the americas so how many lines do we wanna draw here? And why are we drawing these lines?" and they walked away from me.

(laughs)

***** When people cme up and they ask me a question am gonna ask them a question too. hes asking me to do this their maestro is asking meto do this, when i answer the questions. thats what he asked me to do, so thats why  im doing it, you know? Im a guest, can you do this? yeah, yea, ill do it the best i can
so i donttry to... ive never played with a huehuetl withoiut a huehuetl lplayer next to me cause thats ot what i do, so when i play the drum, i if im a, if im asked to give them a demonstration, i will demonstrate the  rhythm as i was thought n the traditional sense, but if im playing non dance rhytms or danza rhytms without a dancer, i will go on another direction... But keeping that as the core rhytm, but ill explain that, say dont think this is what they were doing, thats not theyre doing this instead

Ah cause its very important to me,justlike in india, its very important to me that you let people know but the bad thing is if someone sarches you on youtube they don't know what you said before, or if you said anything, so thats the bad part. Yeah, but the other thing is that ive seen many youtubers where people are talking about, whatever theyre talking about and theyre saying "this goes back thousand, twenty, thirthy thousand years" and you say ah this is... not good.

yeah

so to me its like everybody needs like a basic history lesson

Peiple wanna believe that, ....

Well, when you tell them just ask them "what is theoldest civilizations that we have today" its blabla which is mayan, and 800, and thats between 500 and 800 i think, they dont even have an actual... they say 790,but you now valley in mexico, mexica, it was 300 years later.

So, yeah, its very... and the... like today i was showying my wife if you type  in mayan dance, heyre doing mexican dance and theyre dressed like mexica and its likke... you know? There's no...

and im not saying "it should out of contex", you know, abraham lincoln discovered the united states and then wrote the declaration of independence, no, no, thats wrong, and thats basic stuff

We're not talking about you know, you know his whole arc, just like... no, you got to at least get the order right.

Like one pox did not disscovered to 1947, prior to 1947 the mayan wer thought as these malevolent astronoments,people had no idea that they were doing all these other things and then the pyramids shift it, it changes its like yea! how could it not and now everythings different

So we all dont  know everything, theres all, theres always people that know less than us, always people that know more than us, with be respectful, you dont know who knows what and who doesnt, you dont knwo where theyve been or what they have or what theyve got through, so you have to  say u know based on what i know, this is why i believe this but if you give meinformation that shows thats wrong then im not gonna continue believing the wrog information, i mean...

Or you shouldnt, orri anyway...

yeah, so its really interesting ah, but even, now i know why indians dont write anything down, theres actually a play with a gentleman who ah you kow the indians visited the...so i should tell you,so my teacher was a court musician prior to independence of india, so  that means you have to play secular music, religious music, dance music and cassical music every day for the rag, every day, every day, and ......what he describes the music being played one baptism is....... and all the, all the, its, its, there's, there's written documentation of what it looked like, what it sounded like, how many players and all this kind of stuff. But we still dont know what it sounded like, we wouldn know how the melodies were constructed we dont know any of that

HAve sources have you used in trying to approach begining to think about what aztec music sound like tooday, knowing that it can be the same, obviously....

40: 00
Well i think that, the way its constucted, ah ill give Let me tell you we were in pozos last year, last year? yeah last year, and ah, there was a... we met four instrument makers, and this guy had a room like this big full of drums so im looking at the drums and all of them like theyre tlapanhuehuetsl or tepoz but there different but theyre all tlapanhuehuetl or teponaztli so youre looking at this and we were talking about fluts not being pentatonic, so youre looking at these and none of the flutes have the same tunning even **** so youre looking at these drums so i ask him i said... i felt dumm after i ask that question, i said i have a question for you, e says yeah,, and i said the huehuetls are all different, not the cavin but the high, all that kind of stuff, same with tepos, i said why are they different? why didnt you make 20 of these, cause you could and then he looked at me he smiled and he goes"because everybody is different"

so you can come here and play five of these drums but youre gonna take one, because thats the one that spoke ti you, thats what theres no... thats why i dont make 20 look the same.

Im went to see the flute maker same thing, like this, how can these gentlemants are dos gentlemas o because... because just like youre not her and hses not you, were all different
and the concept of.... time, right? like you know, their calendar, their solar calendar and their lunar calendar, so they obviously new the days days, in their time, right?
and yet, theres no signature on any part of it on any murals on any art on any pyramids, why?! why?

to me its because they were very concerning with now, now only existss now, now is here right now,   theyre not concerned about posterity, they can say ... they did, theres not signature on anything, not even on the codices.

you only... people we know who did, NEtzahualcoyotl, some of the of the kings, the war kings, those guys are named after documentation but everybody else isnt, cause everybody was known by you would just see the thing and say oh thats christina's work,you know? so the thing that now 2019, nobody is here now. music only exists now, when you stop playing its gone, a piece of paper theres no music on this, and people are never here, so thats what interesting about texts, its like, theyre sending you a tex now but your not reading it now, you can read it the day later or four hours later, and whatever it was taht person is not there anymore, theyre not there....so nobody is here anymore, ive noticed, nobody is here...

When i do lectures ah, nah... I was a Coast LA you know? 150 students say 8:30 in the mrning and i said "raise your hand if youre not here" and theyre like... 8laughs) do i raise my hand? like what did you say?  i said raise youre gand if youre not here and then this one lady raised her hand and i said, you could put your hand down, youre not here, she says im not here, i sworn you right now, cause im trying to get to clase, i said ok, i said well theres a 149 that are here, so.... you told me when youre here and then well start, she said, im here now (laiughs)

And i noticed people are not, theyre not

and to me that whole culture is just now. like i said theres no, theres no name on anything, on anything!istn that whats very interesting? Its' very... its like, they were native, its like i wanna be remembered, no...

so my daughter god bless ger, you now my website is called indigenous instruments of mexico and mesoamerica she said dad, your name is not on your website, i go i know, she just wise on the website and  i just, i just for that reason, i said im keeping that alive im not puting my name on it. Shes like "no" but ah... Indigenous, like, if you live in a condimenium in Los Angeles, ok, theres city and theres culture. 

you know? Im irish, my friend from boston, "im irish" hell tell you his grand father, great grandfather, his great great grandfather, all thatstuff i said "have you been to ireland?" yea im "ok", "but im going to ireland" "well call me when you get back i wanna talk to you" so he ok, so he comes back i said "how was it" hes like oh, i went to curk, i went to belfas, i went to i visit all these lovely places, i said i have 3 questions for you did anybody in ireland thought you were ireland? he said no, what do they think you were? an american, he didnt..... I said yeah, do you speak galic? do you wear his clothes? do you eat irish foodo? no. So ethnically youre an irish man, but culturally you dont have anything in common with them anymore...

So,when i hear people say "im italian"or "im escandinavian" or whatever they are its like,,, you know? thats one side of it... the other side of it is, your memory, if you speak to rayman , god bless him, if you ask rayman something, my daughter was a... 8 months old when my father past when he was 90, so she never experienced my father as an adult, and she knows stories about my dad etc etc and they got film with her with my dad etc etc but there's thing that she does that my father does that nobody else in the family does, that shes never experienced that we never talked about, ever, like one thig i cant even do but ah... my father used, ah, it sounds like this 8plays drum), and he's able to do it with hisk nockles, (pplaysdrums) like that... she did it, one day, (mimmics drum), but nobody is ever talked about it, theres no film of him, we never mentioned it... and ah, you know...., its like revueltas, chavez, right? Chavez was like "i wanna find the indigen smell and i wanna oschertrated it" blablabla and i wanna make these things, riht?, and then revuelta is like im not gonna do that, because im mexican, because its in my blaood

and rayman , if you talked to rayman hell do it sometimes, hell say its in the blood, and ah... so playing symphony in india, ah, the conductor.... 

(she speaks)

I remember as youre speaking, ah, he called me and god bless him, he called me and he said "ah! Were gonna do some money in india" and i was shock to find it we dont have any indigenous instruments, but he was hocked to find that? yeah, wehadchines we had african we had japanese indonesian, we have balonese, in ucla, we dont hava a rattle, "we dont have a rattle " he said! (laughs) a one instrument... and i was like... His ah, student, conducting student, i dont know if he asked him to do it or he, i dont know who asked who but he's brazilian but he wanted to do that piece, so he said he talked to steve loza and said u know i dont know where you get this "ill call chris" hes got insturments,so he called me out, he said, he told me..... he said they were wondering if they could i said yeah! I sayd ill tell you what, ah, bring them down to my studio ahm, cause it was supposed to be 4 percusionists, one conductor and the conductor student or so i was told, i said ok, bring them to my studio, well work for two hours on sunday, and then.... ill work with them two more times and the dress rehearsal like 8 isntrumetns when they play the instruments so they can get a feel to that cause they dont responde in any way like a western instrument in any way, so he said, so he calls me back and hes like "theres 16 people coming" jaja like 16 people? you said there were 6 but now other people wanted to come so what i did is i just said the instruments upand then i had them play, and then i did them history and i letthem play and i said, and the conducters are here, and i said to the conducters ah, this is you know = its the conducter medium, their orchestrai sadi this is up too you, do you want this things to stick out or you want them to blend, 

(50:009)

